EDIpack: A generic and interoperable Lanczos-based solver for quantum impurity problems

### 2025-06-01 01_intro.tex
Different colors for different languages

\section{Introduction and Motivation}\label{SecIntro}
Quantum impurity models play a central role in the study of strongly correlated electron systems, capturing the coupling of a small number of localized degrees of freedom to an extended non-interacting environment \cite{Nozieres1980JP,Hewson1993}. 
Originally introduced to describe diluted magnetic impurities in metallic hosts, single-impurity models have led to the full understanding of the Kondo problem \cite{Anderson1961PR,Kondo1964POTP,Schrieffer1966PR} have progressively become powerful tools for exploring a wide range of quantum many-body phenomena \cite{Wilson1975RMP,Georges1996RMP,Kotliar2004PT,Kotliar2006RMP}. More recently, their importance has expanded beyond the realm of condensed matter, influencing the understanding of other areas such as quantum information \cite{Su2013MPLB,Walsh2019PRL,Walsh2020PQ,Walsh2021PNAS,Bellomia2024PRB} or ultra-cold atomic systems \cite{Dao2007PRL,Amaricci2014PRA,Del-Re2018PRA,Walsh2019PRB,Tusi2022NP}.

At the heart of quantum impurity models is the interaction between a localized ``impurity''  and a surrounding bath of itinerant particles, usually represented as a conduction band of electrons \cite{Hewson1993}. In this context, the  impurity  can be defined more generally, so as to include orbital and spin local degrees of freedom as well as a finite number of lattice sites. 
This setup captures many essential aspects of strong correlation, including screening \cite{Roekeghem2014PRL,Roekeghem2014EL,Werner2016JOPCM,Tomczak2017TEPJST}, decoherence and entanglement in a computationally tractable way \cite{Walsh2021PNAS,Bellomia2024PRB}.
Most notably, quantum impurity problems lie at the core of many quantum embedding methods, like density-matrix embedding theory (DMET) \cite{Knizia2012PRL}, Gutzwiller approximation (GA) \cite{Lanata2015PRX,Mejuto-Zaera2023PRB},
Variational Cluster Approximations \cite{Potthoff2003TEPJBCMACS,Senechal2008,Potthoff2011ACP,Nuss2011,Dionne2023SPC,Dionne2023SPCa}
or Dynamical Mean-Field Theory (DMFT) \cite{Georges1996RMP, Kotliar2004PT,Kotliar2006RMP}.
Within these approaches the complex physics of the lattice systems gets reduced to an effective self-consistent quantum impurity problem.
By accounting for local quantum fluctuations, DMFT successfully describes the most important physical features of correlated materials, including Mott insulators, heavy fermion compounds and unconventional superconductors.

>>>>

\section{Introduction and Motivation}\label{SecIntro}
Quantum impurity models play a central role in the study of strongly correlated electron systems, capturing the coupling of a small number of localized degrees of freedom to an extended non-interacting environment \cite{Nozieres1980JP,Hewson1993}. 
Originally introduced to describe diluted magnetic impurities in metallic hosts, single-impurity models have led to the full understanding of the Kondo problem \cite{Anderson1961PR,Kondo1964POTP,Schrieffer1966PR} have progressively become powerful tools for exploring a wide range of quantum many-body phenomena \cite{Wilson1975RMP,Georges1996RMP,Kotliar2004PT,Kotliar2006RMP}. More recently, their importance has expanded beyond the realm of condensed matter, influencing the understanding of other areas such as quantum information \cite{Su2013MPLB,Walsh2019PRL,Walsh2020PQ,Walsh2021PNAS,Stocker2022,Bellomia2024PRB} or ultra-cold atomic systems \cite{Dao2007PRL,Amaricci2014PRA,Del-Re2018PRA,Walsh2019PRB,Tusi2022NP}.

At the heart of quantum impurity models is the interaction between a localized ``impurity''  and a surrounding bath of itinerant particles, usually represented as a conduction band of electrons \cite{Hewson1993}. In this context, the  impurity  can be defined more generally, so as to include orbital and spin local degrees of freedom as well as a finite number of lattice sites. 
This setup captures many essential aspects of strong correlation, including screening \cite{Roekeghem2014PRL,Roekeghem2014EL,Werner2016JOPCM,Tomczak2017TEPJST}, decoherence and entanglement in a computationally tractable way \cite{Walsh2021PNAS,Bellomia2024PRB}.
Most notably, quantum impurity problems lie at the core of many quantum embedding methods, like density-matrix embedding theory (DMET) \cite{Knizia2012PRL}, the Gutzwiller approximation (GA) \cite{Lanata2015PRX,Mejuto-Zaera2023PRB},
Variational Cluster Approximations \cite{Potthoff2003TEPJBCMACS,Senechal2008,Potthoff2011ACP,Nuss2011,Dionne2023SPC,Dionne2023SPCa}
or Dynamical Mean-Field Theory (DMFT) \cite{Georges1996RMP, Kotliar2004PT,Kotliar2006RMP}.
Within these approaches the complex physics of the lattice systems gets reduced to an effective self-consistent quantum impurity problem.
By accounting for local quantum fluctuations, DMFT successfully describes the most important physical features of correlated materials, including Mott insulators, heavy fermion compounds and unconventional superconductors.

### 2025-06-01 05_examples.tex
Different colors for different languages

\subsection{Hubbard model on the Bethe lattice (Fortran API, {\tt ed\_mode=normal})}\label{SecExamplesBetheDMFT}
The description of the Mott transition using the Hubbard model on the Bethe lattice is conventionally regarded as the standard test bed for any DMFT implementation.  \cite{Georges1996RMP,Rozenberg1999PRL,Kotliar1999EPJB,Kotliar2000PRL,Kotliar2002PRL}.

@@

\begin{center}
\begin{minipage}[t]{0.49\linewidth}
\textbf{Fortran}
\begin{lstlisting}[style=fstyle,frame=none,numbers=none,basicstyle={\scriptsize\ttfamily}]
program ed_hm_bethe
   USE EDIPACK
   USE SCIFOR
   implicit none
   integer                :: Le=1000
   real(8)                :: wmixing=0.5d0
   real(8)                :: D=1d0
   integer                :: Nb
   real(8),allocatable    :: Bath(:)
   complex(8),allocatable :: Hloc(:,:,:,:)
   real(8),allocatable    :: Ebands(:),Dbands(:)
   complex(8),allocatable :: Smats(:,:,:,:,:)
   complex(8),allocatable :: Delta(:,:,:,:,:)
   !...  
   !EDIpack: Read variables
   call ed_read_input('inputED.conf')

@@

#Generate Hk=${\color{comment-color}\e(k)}$ and set Hloc
kgrid   = generate_kgrid(Nk)
Hk      = h_square2d(kgrid,t_hop)
Hloc    = np.sum(Hk,axis=0)/Nk**2
ed.set_hloc(Hloc.astype(complex))

>>>>

\subsection{Hubbard model on the Bethe lattice (Fortran/C++ API, {\tt ed\_mode=normal})}\label{SecExamplesBetheDMFT}
The description of the Mott transition using the Hubbard model on the Bethe lattice is conventionally regarded as the standard test bed for any DMFT implementation.  \cite{Georges1996RMP,Rozenberg1999PRL,Kotliar1999EPJB,Kotliar2000PRL,Kotliar2002PRL}.

@@

\begin{center}
\begin{minipage}[t]{0.49\linewidth}
\textbf{Fortran}
\begin{lstlisting}[style=fstyle,frame=none,numbers=none,basicstyle={\scriptsize\ttfamily}]
program ed_hm_bethe
   USE EDIPACK
   USE SCIFOR
   implicit none
   integer               :: Le=1000
   real(8)               :: wmixing=0.5d0
   real(8)               :: D=1d0
   integer               :: Nb
   real(8),allocatable   :: Bath(:)
   complex(8),allocatable:: Hloc(:,:,:,:)
   real(8),allocatable   :: Ebands(:),Dbands(:)
   complex(8),allocatable:: Smats(:,:,:,:,:)
   complex(8),allocatable:: Delta(:,:,:,:,:)
   !...  
   !EDIpack: Read variables
   call ed_read_input('inputED.conf')

@@

#Generate ${\color{comment-color}H_k=\e(k)}$ and set Hloc
kgrid   = generate_kgrid(Nk)
Hk      = h_square2d(kgrid,t_hop)
Hloc    = np.sum(Hk,axis=0)/Nk**2
ed.set_hloc(Hloc.astype(complex))
